Highly Available Transactions: Virtues and Limitations

### 2013-04-01 writeup/evaluation.tex
eval

\begin{figure}
\includegraphics[width=.8\columnwidth]{figs/macrotracelegend.pdf}\vspace{-2mm}
\includegraphics[width=\columnwidth]{figs/lat-thru-trace-K0-SCALEMACRO.pdf}
\includegraphics[width=\columnwidth]{figs/lat-thru-trace-K0-SCALEMACRO.pdf}
\caption{Latency-throughput for database operation on EC2 (DUMMY FIG)}
\label{fig:wan-exp}
\end{figure}

2.) We ran YCSB on EC2 (and maybe TPC-C if we have time) with a few models:
	- Eventual Consistency
	- Eventual consistency with a "master" for updates--simulates the lower bound on a non-HAT system.
	- Transactional Atomicity and Read Committed

	-This is going to look a lot like the ping tests: much higher latency for "non-HATs."
	-Our Naive 2PL implementation bottlenecks extremely quickly.
	-On, say, 5 servers, can get hundreds of thousands of TPS

	-Not meant as an exhaustive study, but validates our intuitions. Coordination costs of non-HAT systems are unaccounted for, will only go up, and we haven't even talked about availability.

>>>>

To better understand the impact of HAT versus non-HAT algorithms and
validate the latency-based analysis in Section~\ref{sec:latency}, we
deployed a basic distributed database prototype on public cloud
infrastructure. We compared the behavior of eventual consistency,
per-key mastering (wherein all operations required a RTT to a
designated master server), Read Committed, and TA with read your
writes guarantees. We deployed an experimental Java-based distributed
datastore backed by LevelDB both within and across EC2 regions and ran
the YCSB benchmark to determine the resulting performance impact. We
partitioned the database into two fully replicated clusters of three
\texttt{m1.xlarge} instances. For HAT, we stuck YCSB clients to their
local cluster and, to provide transactional functionality, grouped
every six YCSB operations from the default workload to form a
transaction. We report the median of five trials per experiment.

\begin{figure}[t!]
\begin{center}
\hspace{2em}\includegraphics[width=.8\columnwidth]{figs/strategylegend.pdf}
\end{center}\vspace{-2em}
\begin{center}\textbf{Within \texttt{us-east}}\end{center}\vspace{-1.5em}
\includegraphics[width=\columnwidth]{figs/lan-threads-lats.pdf}\vspace{-1em}
\includegraphics[width=\columnwidth]{figs/lan-threads-thru.pdf}
\begin{center}\textbf{Between \texttt{us-east} and \texttt{us-west-2}}\end{center}\vspace{-1.5em}
\includegraphics[width=\columnwidth]{figs/wan-threads-lats.pdf}\vspace{-1em}
\includegraphics[width=\columnwidth]{figs/wan-threads-thru.pdf}
\caption{YCSB performance for two clusters of three servers each
  deployed within a single datacenter and cross-datacenters.}
\label{fig:wan-exp}
\end{figure}

As Figure~\ref{fig:wan-exp} shows, within a single datacenter, a
mastered implementation achieves approximately double the latency and
slightly over half of the throughput of a HAT implementation. This is
because, given two replicas for each key, the mastered implementation
serves requests from only one. The mastered implementation still
achieves $19,797$ transactions per second---over 6,000 per
master---and is only $15.56$ms slower at peak compared to eventual
consistency. However, eventual consistency and HAT models achieve
substantially higher throughput--$37,575$ and $36,178$ transactions
per second, respectively--and lower latency---$48.3\%$ and $47.2\%$
respectively. This difference is unsurprising because, in these weaker
models, all servers can serve both reads and writes, effectively
doubling the capacity of the system. The weaker models still propagate
new values between replicas, but this process is asynchronous and,
aside from consuming disk I/O and network bandwidth, does not
substantially impact clients. As expected, HAT models perform
similarly to eventual consistency: our Read Committed implementation
buffers client writes and adds no additional server-side overhead,
while TA requires additional metadata with each write but has less
than 5\% performance impact compared to eventual consistency.

In comparison, over multiple datacenters, the cost of coordination
increases. The HAT and eventual consistency implementations continue
to enjoy local RTTs only, and, with the exception of a slight
performance degradation in TA---which we believe is attributable to
Java garbage collection overheads in TA's inbound anti-entropy
buffers---performance remains unchanged. However, each YCSB update to
a remote replica ($\frac{2}{3}$ in one datacenter, $\frac{1}{3}$ in
the other) of requiring a remote RTT between Virginia and Oregon (or
vice-versa) to reach the master. This leads to a latency increase of
around $166$ms ($746\%$) per transaction compared to a
single-datacenter deployment. In order to attain reasonable throughput
for the mastered deployment, we had to greatly increase the number of
concurrent clients accessing the database, and we found that threading
and socket overheads quickly overwhelmed the database server. Over a
wide-area network, the increased cost of RTTs dramatically affects
system efficiency.

This brief set of experiments is not meant as an exhaustive study, but
it validates our earlier intuition regarding the cost of violating
HAT-compliance. As Deutsch points out, ignoring factors such as
latency can ``cause big trouble and painful learning
experiences''~\cite{fallacies-deutsch}---in a single-site context,
paying the cost of coordination may be tenable, but, especially as
services are geo-replicated, the costs increase.
